### 2021-07-30 Project 3/Assignment/Creditworthiness/Submission - NND.Rmd
Revise submission file

---
title: 'Project Creditworthiness'
author: "Dung Nguyen"
output: 
  pdf_document:
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error=TRUE)
```

## Step 1: Business and Data Understanding
###	What decisions needs to be made?
Determining whether customers are creditworthy to give a loan to.

###	What data is needed to inform those decisions?
- Data set of past applications for training models:
  - Outcome: Credit-Application-Result.
  - Potention predictors: Duration_of_Credit_Month, Credit_Amount, Age_years, Instalment_per_cent, Most_valuable_available_asset, Type_of_apartment, and so on.
- New data set (without outcome variable) to apply the chosen model and make decisions.

###	What kind of model (Continuous, Binary, Non-Binary, Time-Series) do we need to use to help make these decisions?
Binary model.

## Step 2: Building the Training Set
After importing, the data set were formatted as suggested.
```{r setwd, echo=FALSE}
setwd("C:/Users/ngocdung/Dropbox/Front-end developer/src/data-science-path/udacity-nanodegree/Project 3/Assignment")
raw <- read.csv("./Creditworthiness/credit-data-training.csv", stringsAsFactors=F)
validation <- read.csv("./Creditworthiness/customers-to-score.csv", stringsAsFactors=F)
```

```{r import, message=c(32), warning=FALSE}
library(tidyverse)
library(caret)
library(randomForest)
library(bst)
library(cowplot)
library(pROC)
#raw <- read.csv("./Creditworthiness/credit-data-training.csv", stringsAsFactors=F)
#validation <- read.csv("./Creditworthiness/customers-to-score.csv", stringsAsFactors=F)
colnames(validation) <- str_replace_all(colnames(validation), '\\.', '_')
colnames(raw) <- str_replace_all(colnames(raw), '\\.', '_')
glimpse(raw)

```

### Identifying and imputing missing data
The number of missing values in each field:
```{r missing}
isna = vector(length = length(raw))
i=1
for(col in raw){
  isna[i] = sum(is.na(col))
  i = i + 1 }
print(data.frame(field = colnames(raw), no_of_missing = isna))
```

`Duration_in_Current_address` has 344 missing values so I dropped it.
`Age_years` has 12 missing values so I imputed it by median.
```{r impute and drop}
raw = raw[,!(colnames(raw) %in% c('Duration_in_Current_address'))]

median_age <- median(raw$Age_years[!is.na(raw$Age_years)])
raw$Age_years[is.na(raw$Age_years)] <- median_age
mean(raw$Age_years)
```

### Identifying low-variability fields and removing them
Frequency table and histogram of low-variability fields:
```{r drop}
print(list(count(raw, Occupation), 
     count(raw, No_of_dependents),
     count(raw, Foreign_Worker),
     count(raw, Guarantors),
     count(raw, Concurrent_Credits)))

p1 <- histogram(raw$Occupation)
p2 <- histogram(raw$No_of_dependents)
p3 <- histogram(raw$Foreign_Worker)
p4 <- histogram(as.factor(raw$Guarantors))
p5 <- histogram(as.factor(raw$Concurrent_Credits))
plot_grid(p1, p2, p3, p4, p5)
```

After removing those fields plus `Telephone`, no high correlation was detected among numeric fields:
```{r correlation}
numeric <-  c('Duration_of_Credit_Month', 
              'Credit_Amount', 
              'Age_years')
numeric2 <- c('Instalment_per_cent',   
           'Most_valuable_available_asset',
           'Type_of_apartment')
factor <- c("Credit_Application_Result",
             "Account_Balance",
             "Payment_Status_of_Previous_Credit",
             "Purpose",
             "Value_Savings_Stocks",
             "Length_of_current_employment",
             "No_of_Credits_at_this_Bank")

raw <- data.frame(raw[c(numeric, numeric2)],
                     apply(raw[factor],2, as.factor))

print(cor(raw[c(numeric, numeric2)]))
```

The final data set has 13 columns
```{r final}
raw <- raw %>% mutate(Credit_Application_Result= as.factor(Credit_Application_Result),
                           Account_Balance= as.factor(Account_Balance),
                           Payment_Status_of_Previous_Credit=as.factor(Payment_Status_of_Previous_Credit),
                           Purpose=as.factor(Purpose),
                           Value_Savings_Stocks=as.factor(Value_Savings_Stocks),
                           Length_of_current_employment=as.factor(Length_of_current_employment),
                           No_of_Credits_at_this_Bank=as.factor(No_of_Credits_at_this_Bank))
data <- raw
glimpse(data) 
```

## Step 3: Train your Classification Models
### Create Estimation and Validation samples
```{r train test set}
data$worthy <- as.factor(ifelse(data$Credit_Application_Result == 'Creditworthy', 1, 0))
data$nonworthy <- as.factor(ifelse(data$Credit_Application_Result == 'Non-Creditworthy', 1, 0))

set.seed(1)
train.index <- createDataPartition(data$Credit_Application_Result, p = .7, list = FALSE)  
train <- data[train.index,]
test <- data[-train.index,] 
```

### Confusion matrices
```{r training}
# Logistic regression
lr <- vector()
model1 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method='glm')  #check
y_hat1 <- predict(model1, test)
y_hat10 <- predict(model1, train)
lr['Accuracy_train'] <- confusionMatrix(data = y_hat10, reference = train$Credit_Application_Result)$overall["Accuracy"] #Bias
lr['Accuracy_test'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$overall["Accuracy"]
lr['PPV'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$byClass['Pos Pred Value']
lr['NPV'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$byClass['Neg Pred Value']

model1w <- train(worthy~., 
                  data=train[!colnames(train) %in% c('Credit_Application_Result', 'nonworthy')], 
                  method='glm')
y_hat1w <- predict(model1w, test)
lr['Accuracy_worthy'] <- confusionMatrix(data = y_hat1w, reference = test$worthy)$overall["Accuracy"]

model1n <- train(nonworthy~., 
                  data=train[!colnames(train) %in% c('Credit_Application_Result', 'worthy')], 
                  method='glm')
y_hat1n <- predict(model1n, test)
lr['Accuracy_nonworthy'] <- confusionMatrix(data = y_hat1n, reference = test$nonworthy)$overall["Accuracy"]

vi1 <- varImp(model1, scale = F)   #pvalue
plot1 <- ggplot(vi1, top = dim(vi1$importance)[1]) + 
  ggtitle('Variable Importance - Logistic Regression')

# Decision tree
tree <- vector()
model2 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')],
                method='rpart')
y_hat2 <- predict(model2, test)
y_hat20 <- predict(model2, train)
tree['Accuracy_train'] <- confusionMatrix(data = y_hat20, reference = train$Credit_Application_Result)$overall["Accuracy"] 
tree['Accuracy_test'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$overall["Accuracy"]            
tree['PPV'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$byClass['Pos Pred Value']
tree['NPV'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$byClass['Neg Pred Value']

model2w <- train(worthy~., 
                 data=train[!colnames(train) %in% c('Credit_Application_Result', 'nonworthy')], 
                 method='rpart')
y_hat2w <- predict(model2w, test)
tree['Accuracy_worthy'] <- confusionMatrix(data = y_hat2w, reference = test$worthy)$overall["Accuracy"]

model2n <- train(nonworthy~., 
                 data=train[!colnames(train) %in% c('Credit_Application_Result', 'worthy')], 
                 method='rpart')
y_hat2n <- predict(model2n, test)
tree['Accuracy_nonworthy'] <- confusionMatrix(data = y_hat2n, reference = test$nonworthy)$overall["Accuracy"]

vi2 <- varImp(model2, scale = F)   
plot2 <- ggplot(vi2, top = dim(vi2$importance)[1]) + 
  ggtitle('Variable Importance - Decision Tree')

# Forest model
forest <- vector()
set.seed(1)
model3 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method='rf',
                ntree=500)
y_hat3 <- predict(model3, test)
y_hat30 <- predict(model3, train)
forest['Accuracy_train'] <- confusionMatrix(data = y_hat30, reference = train$Credit_Application_Result)$overall["Accuracy"]
forest['Accuracy_test'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$overall["Accuracy"]   
forest['PPV'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$byClass['Pos Pred Value']
forest['NPV'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$byClass['Neg Pred Value']

set.seed(1)
model3w <- train(worthy~., 
                 data=train[!colnames(train) %in% c('Credit_Application_Result', 'nonworthy')], 
                 method='rf',
                 ntree=500)
y_hat3w <- predict(model3w, test)
forest['Accuracy_worthy'] <- confusionMatrix(data = y_hat3w, reference = test$worthy)$overall["Accuracy"]

set.seed(1)
model3n <- train(nonworthy~., 
                 data=train[!colnames(train) %in% c('Credit_Application_Result', 'worthy')], 
                 method='rf',
                 ntree=500)
y_hat3n <- predict(model3n, test)
forest['Accuracy_nonworthy'] <- confusionMatrix(data = y_hat3n, reference = test$nonworthy)$overall["Accuracy"]

vi3 <- varImp(model3, scale = F)   #pvalue
plot3 <- ggplot(vi3, top = dim(vi3$importance)[1]) + 
  ggtitle('Variable Importance - Random Forest')

# Boosted model - Boosted Classification Trees
boosted <- vector()
# Data have to be re-formatted for bst() function
train4 <- train %>% mutate(#Credit_Application_Result= as.numeric(Credit_Application_Result),
  Account_Balance= as.factor(as.numeric(Account_Balance)),
  Payment_Status_of_Previous_Credit=as.factor(as.numeric(Payment_Status_of_Previous_Credit)),
  Purpose=as.factor(as.numeric(Purpose)),
  Value_Savings_Stocks=as.factor(as.numeric(Value_Savings_Stocks)),
  Length_of_current_employment=as.factor(as.numeric(Length_of_current_employment)),
  No_of_Credits_at_this_Bank=as.factor(as.numeric(No_of_Credits_at_this_Bank)),
  )

test4 <- test %>% mutate(#Credit_Application_Result= as.numeric(Credit_Application_Result),
  Account_Balance= as.factor(as.numeric(Account_Balance)),
  Payment_Status_of_Previous_Credit=as.factor(as.numeric(Payment_Status_of_Previous_Credit)),
  Purpose=as.factor(as.numeric(Purpose)),
  Value_Savings_Stocks=as.factor(as.numeric(Value_Savings_Stocks)),
  Length_of_current_employment=as.factor(as.numeric(Length_of_current_employment)),
  No_of_Credits_at_this_Bank=as.factor(as.numeric(No_of_Credits_at_this_Bank)),
  )
control <- trainControl(method = "cv", number = 5)
set.seed(1)
model4 <- train(Credit_Application_Result~., 
                data=train4[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method="bstTree", 
                trControl = control)
y_hat4 <- predict(model4, test4)
y_hat40 <- predict(model4, train4)
boosted['Accuracy_train'] <- confusionMatrix(data = y_hat40, reference = train4$Credit_Application_Result)$overall["Accuracy"]  
boosted['Accuracy_test'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$overall["Accuracy"]  
boosted['PPV'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$byClass['Pos Pred Value']
boosted['NPV'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$byClass['Neg Pred Value']

set.seed(1)
model4w <- train(worthy~., 
                 data=train4[!colnames(train) %in% c('Credit_Application_Result', 'nonworthy')], 
                 method='bstTree',
                 trControl = control)
y_hat4w <- predict(model4w, test4)
boosted['Accuracy_worthy'] <- confusionMatrix(data = y_hat4w, reference = test4$worthy)$overall["Accuracy"]

set.seed(1)
model4n <- train(nonworthy~., 
                 data=train4[!colnames(train) %in% c('Credit_Application_Result', 'worthy')], 
                 method='bstTree',
                 trControl = control)
y_hat4n <- predict(model4n, test4)
boosted['Accuracy_nonworthy'] <- confusionMatrix(data = y_hat4n, reference = test4$nonworthy)$overall["Accuracy"]

vi4 <- varImp(model4, scale = F)   #pvalue
plot4 <- ggplot(vi4, top = dim(vi4$importance)[1]) + 
  ggtitle('Variable Importance - Boosted Tree')

#Print confusion matrices
print(confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result))
```
### Overall accuracy
Among the 4 models, Logistic regression has the highest overall accuracy agains the Validation set (0.83). The bias was not observed in Logistic regression, Random forest, and Boosted tree model considering PPV and NPV. 
```{r acc train test}
print(data.frame(lr, tree, forest, boosted)[c('Accuracy_train','Accuracy_test','PPV','NPV'),])
```
### Plot variable importance
Most important predictors:
- For Logistic regression: Account_Balance, Payment_Status_of_Previous_Credit, Credit_Amoun, Value_Savings_Stocks, Purpose, Instalment_per_cent.
- For Decision tree: Payment_Status_of_Previous_Credit, Credit_Amount, Value_Savings_Stocks, Account_Balance, Duration_of_Credit_Month.
- For Random forest: Credit_Amount, Age_years, Duration_of_Credit_Month.
- For Boosted tree: Account_Balance, Value_Savings_Stocks, Paymen_Status_of_Previous_Credit, and Duration_of_Credit_Month are the 4 most important variables. The difference is not remarkable among the fields.
```{r plot1}
plot1
plot2
plot3
plot4
```

## Step 4: Writeup
Accuracy table
```{r acc}
print(data.frame(lr, tree, forest, boosted)[c('Accuracy_train','Accuracy_test','PPV','NPV'),])
```

Plot ROCs 
```{r plot2}
par(pty = "s")
roc(as.numeric(test$Credit_Application_Result), 
    as.numeric(y_hat1), plot=TRUE,
    legacy.axes=TRUE, percent=TRUE, 
    xlab="False Positive Percentage", 
    ylab="True Postive Percentage", 
    print.auc=TRUE,
    print.auc.x=45,
    ) 
legend("bottomright", legend=c("Logisitic Regression"))

plot.roc(as.numeric(test$Credit_Application_Result),
         as.numeric(y_hat2),
         legacy.axes=TRUE, percent=TRUE,
         xlab="False Positive Percentage",
         ylab="True Postive Percentage",
         print.auc=TRUE,
         print.auc.x=45,
)
legend("bottomright", legend=c("Decision Tree"))

plot.roc(as.numeric(test$Credit_Application_Result),
         as.numeric(y_hat3),
         legacy.axes=TRUE, percent=TRUE,
         xlab="False Positive Percentage",
         ylab="True Postive Percentage",
         print.auc=TRUE,
         print.auc.x=45,
)
legend("bottomright", legend=c("Random Forest"))

plot.roc(as.numeric(test$Credit_Application_Result),
         as.numeric(y_hat4),
         legacy.axes=TRUE, percent=TRUE,
         xlab="False Positive Percentage",
         ylab="True Postive Percentage",
         print.auc=TRUE,
         print.auc.x=45,
)
legend("bottomright", legend=c("Boosted Tree"))

```
Logistic regresson model is not bias and has the highest overall accuracy against the Validation set. It is also has the most optimal ROC. Therefore, I choose to use logistic regression model.
Applying the chosen model, the number of individuals are creditworthy is 413.
```{r predict}
yhat_final <- predict(model1, validation)
print(sum(yhat_final=='Creditworthy'))
```